### 2025-07-06 AoP/resubmit-cover-letter.tex
T3: Complete AOP manuscript resubmission process with author name change coordination
- Created professional cover letter with optimized single-page formatting
- Coordinated author name change (Luigi → Lin) through Elsevier's inclusive policy
- Successfully resubmitted AOP manuscript 81029 to Annals of Physics
- Documented complete resubmission process in memory bank system
- Updated task registry with T3 completion and session tracking

\documentclass[11pt]{article}
\usepackage[margin=1.2in,scale=1.2]{geometry}
\usepackage{amsmath}
\usepackage{natbib}
\usepackage{enumitem}
\usepackage{titlesec}

\usepackage{parskip}

\setlength{\parskip}{8pt}
\setlength{\parindent}{0pt}

% Reduce space around sections
\titlespacing*{\section}{0pt}{0.5ex}{0.5ex}
\titlespacing*{\subsection}{0pt}{0.5ex}{0.5ex}

% Reduce list spacing
\setlist{itemsep=0pt, parsep=0pt, topsep=2pt}

\setlength{\baselineskip}{12pt}

\begin{document}

\begin{center}
\textbf{\large Cover Letter for Revised Manuscript Submission}
\end{center}

\vspace{0.2cm}

\noindent
\begin{tabular}{@{}ll}
\textbf{Manuscript Reference Number:} & AOP 81029 \\
\textbf{Journal:} & Annals of Physics \\
\textbf{Date:} & \today \\
\end{tabular}

\vspace{0.3cm}

\noindent
Dear Editor,

We are pleased to submit the revised version of our manuscript entitled ``A Loop Quantum Gravity Inspired Action for the Bosonic String and Emergent Dimensions at Large Scales'' (Reference Number: AOP 81029) for publication in Annals of Physics.

We have carefully addressed all comments and suggestions from the reviewers. A detailed point-by-point response is provided in the accompanying file ``response-to-referee.pdf''.

\textbf{Author Name Change:} One co-author has requested a name change from Luigi Teixeira de Sousa to Lin Teixeira de Sousa. We have contacted namechange@elsevier.com per Elsevier's inclusive author name change policy. This correction does not affect authorship contributions or order.

\textbf{Summary of Revisions:} Major improvements include: (1) Enhanced theoretical foundation with extensive discussions connecting our minimal area approach to existing minimal length proposals in string theory; (2) Clearer physical justification for the expansion parameter $k \ll 1$ and strengthened analogy with T/S-dualities; (3) Clarified anomaly-freedom claims, explaining this refers specifically to absence of central charge anomalies due to finite mode truncation; (4) Added important caveats about approach limitations; (5) Improved interpretation of emergent dimensions and their connection to holographic principles and cosmological horizon scale.

We confirm that: (i) all authors have reviewed and approved the revised manuscript; (ii) the manuscript has not been submitted elsewhere; (iii) all ethical guidelines have been followed; (iv) all necessary permissions have been obtained for copyrighted material.

We thank the reviewers for their constructive feedback and look forward to your favorable consideration.

\vspace{0.3cm}

\noindent
Sincerely,

\vspace{0.4cm}

\noindent
\begin{tabular}{@{}l}
Deepak Vaid \\
502, Broadway Apartments \\
Mangaluru, Karnataka, India \\
Email: dvaid79@gmail.com \\
\textbf{Corresponding Author}
\end{tabular}

\vspace{0.3cm}

\noindent
\textbf{Attachments:} \texttt{response-to-referee.pdf} (detailed response), \texttt{resubmit-cover-letter.pdf} (cover letter), \texttt{lqg-strings.pdf} (revised manuscript), \texttt{lqg-strings-diff.pdf} (diff file), \texttt{lqg-strings.bib} \& \texttt{lqg-strings.bbl} (bibliography files).

\end{document}