### 2025-09-07 AoP/resubmit-cover-letter-r2.tex
Complete T4: Submit final AOP manuscript revisions
- Enhanced Kalb-Ramond field derivation with step-by-step exposition
- Fixed equation label references for improved clarity
- Addressed all referee clarifications in response document
- Successfully submitted revised manuscript for final acceptance
- Updated memory bank

\documentclass[11pt]{article}
\usepackage[margin=1.2in,scale=1.2]{geometry}
\usepackage{amsmath}
\usepackage{natbib}
\usepackage{enumitem}
\usepackage{titlesec}

\usepackage{parskip}

\setlength{\parskip}{8pt}
\setlength{\parindent}{0pt}

% Reduce space around sections
\titlespacing*{\section}{0pt}{0.5ex}{0.5ex}
\titlespacing*{\subsection}{0pt}{0.5ex}{0.5ex}

% Reduce list spacing
\setlist{itemsep=0pt, parsep=0pt, topsep=2pt}

\setlength{\baselineskip}{12pt}

\begin{document}

\begin{center}
\textbf{\large Cover Letter for 2nd Manuscript Revision}
\end{center}

\vspace{0.2cm}

\noindent
\begin{tabular}{@{}ll}
\textbf{Manuscript Reference Number:} & AOP 81029 \\
\textbf{Journal:} & Annals of Physics \\
\textbf{Date:} & \today \\
\end{tabular}

\vspace{0.3cm}

\noindent
Dear Editor,

We thank you for your provisional acceptance of our paper in your esteemed journal.

We have carefully addressed all comments and suggestions from the second round of referee comments. A detailed point-by-point response is provided in the accompanying file ``response-to-referee-r2.pdf''. We hope that we have satisfactorily addresses any concerns the referee might have had.

We confirm that: (i) all authors have reviewed and approved the revised manuscript; (ii) the manuscript has not been submitted elsewhere; (iii) all ethical guidelines have been followed; (iv) all necessary permissions have been obtained for copyrighted material.

We thank the referee and the editor once again for their patient and constructive feedback through this process.

\vspace{0.3cm}

\noindent
Sincerely,

\vspace{0.4cm}

\noindent
\begin{tabular}{@{}l}
Deepak Vaid \\
502, Broadway Apartments \\
Mangaluru, Karnataka, India \\
Email: dvaid79@gmail.com \\
\textbf{Corresponding Author}
\end{tabular}

\vspace{0.3cm}

\noindent
\textbf{Attachments:} \texttt{response-to-referee-r2.pdf} (detailed response), \texttt{resubmit-cover-letter-r2.pdf} (cover letter), \texttt{lqg-strings.pdf} (revised manuscript).

\end{document}